---
title: "Sleep/Productivity A1"
author: "Anthony Fernandez"
date: "2025-01-30"
output: html_document
---

```{r message=FALSE, warning=FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(lmtest)
library(car)
library(ggpubr)

# Read the data
df <- read.csv("Personal Metric Template.csv")

# View the structure of the dataset
str(df)

```

```{r}

df <- na.omit(df) 
df$Date <- as.Date(df$Date, format="%m/%d/%Y")
df$Weekday <- as.factor(weekdays(df$Date))  # Convert to factor



model <- lm(Wake.Up ~ Productivity..Cal. + Sleep.Hours + Stress + Steps + Weekday , data=df)

summary(model)

```

```{r message=FALSE, warning=FALSE}
# Load additional library
library(ggcorrplot)

# Compute correlation matrix (excluding non-numeric columns)
cor_matrix <- cor(df[, c("Wake.Up", "Sleep.Hours", "Stress", "Steps", "Productivity..Cal.")], use="complete.obs")

# Plot correlation heatmap with correct method
ggcorrplot(cor_matrix, method="square", type="lower", lab=TRUE, lab_size=3, colors = c("red", "white", "blue")) +
  ggtitle("Correlation Heatmap of Key Metrics") +
  theme_minimal()

```

```{r}
ggplot(df, aes(x=Weekday, y=Productivity..Cal., fill=Weekday)) +
  geom_boxplot(alpha=0.7) +
  scale_fill_brewer(palette="Set1") +  # Use a colorful palette
  labs(title="Productivity Across Different Days of the Week",
       x="Day of the Week",
       y="Productivity (Cal)") +
  theme_minimal() +
  theme(legend.position="none")  # Remove legend for clarity
```

```{r}
ggplot(df, aes(x=Weekday, y=Sleep.Hours, fill=Weekday)) +
  geom_boxplot(alpha=0.7) +
  scale_fill_brewer(palette="Set1") +  # Use a colorful palette
  labs(title="Sleep Hours Across Different Days of the Week",
       x="Day of the Week",
       y="Sleep Hours") +
  theme_minimal() +
  theme(legend.position="none") 
```



```{r}
ggplot(df, aes(x=Wake.Up, y=Productivity..Cal.)) +
  geom_point(color="orange", alpha=0.7) +  # Scatter points
  geom_smooth(method="lm", color="orange", fill="orange", alpha=0.2) +  # Regression line
  labs(title="Sleep Quality vs. Productivity",
       x="Sleep Quality (Wake Up Score)",
       y="Productivity (Cal)") +
  theme_minimal()
```

```{r}
ggplot(df, aes(x=Wake.Up, y=Stress)) +
  geom_point(color="red", alpha=0.7) +  # Scatter points
  geom_smooth(method="lm", color="red", fill="red", alpha=0.2) +  # Regression line
  labs(title="Sleep Quality vs. Stress",
       x="Sleep Quality (Wake Up Score)",
       y="Stress Score") +
  theme_minimal()
```


```{r warning=FALSE}
cor_value <- cor(df$Stress, df$Wake.Up, use = "complete.obs")


ggplot(df, aes(x = Stress, y = Wake.Up, color = Stress, size = Sleep.Hours)) +
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm", color = "black", linetype = "dashed", se = FALSE) +
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "How Stress Affects Wake Up Score",
       subtitle = paste("Correlation:", round(cor_value, 2)),
       x = "Stress Level",
       y = "Wake Up Score",
       color = "Stress Level",
       size = "Sleep Hours") +
  theme_minimal() +
  theme(legend.position = "right")
```



```{r}
ggplot(df, aes(x=Weekday, y=Stress, fill=Weekday)) +
  geom_boxplot(alpha=0.7) +
  scale_fill_brewer(palette="Set1") +  # Use a colorful palette
  labs(title="Stress Across Different Days of the Week",
       x="Day of the Week",
       y="Stress") +
  theme_minimal() +
  theme(legend.position="none")
```






My findings from this data show that better sleep quality tends to lead to higher productivity, especially on weekdays. The scatter plot suggests that higher wake-up scores correlate with burning more calories, meaning when sleep is better, I am more productive. Interestingly, productivity levels seem to be higher on weekdays compared to weekends, my guess is that this is due to more structured routines during the week. There’s also a negative relationship between stress and productivity, and stress levels might be lower on weekends, but so is productivity. Another interesting trend is that taking more steps throughout the day is linked to better productivity, so staying active might help performance (calories burned) no matter what day it is.